---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# qboost

<!-- badges: start -->
[![R-CMD-check](https://github.com/SvenKlaassen/qboost/workflows/R-CMD-check/badge.svg)](https://github.com/SvenKlaassen/qboost/actions)
[![Codecov test coverage](https://codecov.io/gh/SvenKlaassen/qboost/branch/master/graph/badge.svg)](https://codecov.io/gh/SvenKlaassen/qboost?branch=master)
<!-- badges: end -->

High-dimensional quantile regression using boosting. This repository is the official implementation of ...

Still in progress!

## Description

Quantile Regression is a popular tool to model the conditional quantiles of a response variable. In a high-dimensional setting, the theoretical results are largely limited to a penalized quantile regression model([Belloni and Chernozhukov, 2011](https://projecteuclid.org/journals/annals-of-statistics/volume-39/issue-1/%e2%84%931-penalized-quantile-regression-in-high-dimensional-sparse-models/10.1214/10-AOS827.full)).<br />
Instead, we employ a greedy procedure to gradually improve the regression function. To overcome the non-differentiability of the check loss funciton, we employ the convolution-type smoothed quantile regression ([Fernandes, Guerre and Horta, 2019](https://www.tandfonline.com/doi/abs/10.1080/07350015.2019.1660177?journalCode=ubes20) and [He, Pan, Tan, Zhou](https://arxiv.org/abs/2012.05187)). 

## Installation qboost

To install development version from [GitHub](https://github.com/), run the following commands:

``` r
# install.packages("devtools")
remotes::install_github("SvenKlaassen/qboost")
```

## Examples

A simple example. We generate data from a basic linear model 
$$ Y = \beta_0 + X^T\beta +\epsilon,$$
where $\beta_0=1$ and $\epsilon\sim t_2$. To account for a high-dimensional setting, we only generate $n = 200$ observations, whereas the covariates $X$ are generated from a standard multivariate gaussian distribution of dimension $p = 500$. The coefficient vector is set to $\beta_j = 1$ for the first $j = 1,\dots s$ components and $0$ elsewhere. 

```{r dgp}
library(MASS)
set.seed(42)
n <- 200; p <- 500; s <- 4
beta <- rep(c(1,0),c(s+1,p-s))

X = mvrnorm(n, rep(0, p), diag(p))
epsilon = rt(n, 2)
Y = cbind(1, X) %*% beta + epsilon
```

Next, we use the orthogonal variant of the qboost algorithm (WCGA, by setting `stepsize = NULL`) and proceed with for $10$ greedy selection steps.


```{r example}
library(qboost)
n_steps <- 199; tau <- .5
model_WCGA <- qboost(X,Y, tau = tau, m_stop = n_steps, h = 0.2, kernel = "Gaussian", stepsize = NULL)
# selected covariates
print(model_WCGA$selection_path[1:10])
```
Additionally, we employ the Weak Restricted Greedy Algorithm.
```{r example2}
model_WRGA <- qboost(X,Y, tau = tau, m_stop = n_steps, h = 0.2, kernel = "Gaussian", stepsize = 0.1)
```


To compare the performance to the oracle model, we apply the conquer algorithm to the first $s$ components of $X$ and report the norm of the coefficient vectors.

```{r exampleplot, warning=FALSE}
library(conquer)
fit.conquer <- conquer(X[,1:s], Y, tau = tau, h = 0.2, kernel = "Gaussian")
norm_conquer <- sqrt(sum((fit.conquer$coeff-beta[1:(s+1)])^2))

library(ggplot2)
library(reshape2)
df <- melt(data.frame("norm_WCGA" = apply(model_WCGA$coeff_path,2,
                                       function(x) sqrt(sum((x-beta)^2))),
                 "norm_WRGA" = apply(model_WRGA$coeff_path,2,
                                       function(x) sqrt(sum((x-beta)^2))),
                 "step" = 0:n_steps),
            id = c("step"))
        

ggplot(df) +
  geom_line(aes(x = step,y = value, color = variable), size = 1) + 
  scale_colour_manual(values=c("red","blue")) +
  geom_hline(yintercept = norm_conquer, color = "Forestgreen", size = 1) +
  ylim(c(0,2.5)) +
  labs(title = "Norm of the estimators with increasing stepsize",
       x = "Steps",
       y = "Norm",
       colour = "Algorithm") +
  theme(legend.position = 'bottom')
```


## References
* He, Xuming, et al. "Smoothed quantile regression with large-scale inference." arXiv preprint arXiv:2012.05187 (2020). [Paper](https://arxiv.org/abs/2012.05187)

* Belloni, Alexandre, and Victor Chernozhukov. "ℓ1-penalized quantile regression in high-dimensional sparse models." The Annals of Statistics 39.1 (2011): 82-130. [Paper](https://projecteuclid.org/journals/annals-of-statistics/volume-39/issue-1/%e2%84%931-penalized-quantile-regression-in-high-dimensional-sparse-models/10.1214/10-AOS827.full)

* Fernandes, Marcelo, Emmanuel Guerre, and Eduardo Horta. "Smoothing quantile regressions." Journal of Business & Economic Statistics 39.1 (2021): 338-357. [Paper](https://www.tandfonline.com/doi/full/10.1080/07350015.2019.1660177?casa_token=bkJ73Q8oXYIAAAAA%3A0g8P9Bb5elGlCBU8_bsuN_oLauFgMfQcojZmI4ERJO1WVD1M5UOw1RLRix7mMxDMMWukfZR-sbE)

