---
title: "Introduction to Quantile Boosting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro_qboost}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

knitr::opts_chunk$set(
 fig.width = 6,
 fig.asp = 0.8,
 out.width = "80%",
 fig.align = "center"
)
```

High-dimensional quantile regression using boosting. This repository is the official implementation of ...

**TODO: Add Article from arXiv when available**

Quantile Regression is a popular tool to model the conditional quantiles of a response variable. In a high-dimensional setting, the theoretical results are largely limited to a penalized quantile regression model ([Belloni and Chernozhukov, 2011](https://projecteuclid.org/journals/annals-of-statistics/volume-39/issue-1/%e2%84%931-penalized-quantile-regression-in-high-dimensional-sparse-models/10.1214/10-AOS827.full)).<br />
Instead, we employ a greedy procedure to gradually improve the regression function. To overcome the non-differentiability of the check loss funciton, we employ the convolution-type smoothed quantile regression ([Fernandes, Guerre and Horta, 2019](https://www.tandfonline.com/doi/abs/10.1080/07350015.2019.1660177?journalCode=ubes20) and [He, Pan, Tan, Zhou](https://arxiv.org/abs/2012.05187)). 

## Installation qboost

To install development version from [GitHub](https://github.com/), run the following commands:

``` r
# install.packages("devtools")
remotes::install_github("SvenKlaassen/qboost")
```

## Introduction to Quantile Regression

In quantile regression the conditional $\tau$-quantile of some response variable $Y$ is fitted. Here, we assume the conditional quantile 
$$F_{Y|X}^{-1}(\tau|x)=x^T\beta_\tau$$
to be a linear function of the covariates. It can be shown
$$\beta_\tau = \arg\min_{\beta\in\mathbb{R}^p}\mathbb{E}[\rho_\tau(Y - X^T\beta)],$$
where 
$$\rho_\tau(x) = (\tau - I_{\{x\le 0\}})x$$
is the check function ([Quantile Regression](https://en.wikipedia.org/wiki/Quantile_regression)).

```{r fig_loss, echo=FALSE}
library(qboost)
library(ggplot2)
x <- seq(-1,1,0.01)
tau_vec <- c(0.25,0.5,0.75)
loss <- lapply(tau_vec, function(tau) (tau - (x <= 0))*x)
df <- data.frame("Loss" = unlist(loss),
                 "Quantile" = rep(c("0.25","0.5","0.75"),each = length(x)),
                 "x" = rep(x,3))

ggplot(data = df, aes(x = x, y = Loss, color = Quantile)) + 
  geom_line() + 
  geom_hline(yintercept = 0)
```
Since the loss function is nondifferentiable at zero, the theoretical results rely on a smoothed version of the loss, based on a kernel density estimator.
Given a symmetric kernel $k:\mathbb{R}\to [0,\infty)$ and a bandwidth $h$, the smoothed loss is defined as the convolution
$$\rho_{\tau,h}(x):=(\rho_\tau*k_h)(x)=\int_{-\infty}^\infty\rho_\tau(y)k_h(y-x)dy$$
with $k_h(x)=(1/h)k(x/h)$.

In the plot you can see different examples using different kernels (all with bandwidth $h = 0.1$).
```{r fig_smoothed_loss, echo=FALSE}
h <- 0.1
loss_gaussian <- lapply(tau_vec, function(tau){
  l = (2/pi)^(1/2)*exp(-(x/h)^2/2)+(x/h)*(1-2*stats::pnorm(-(x/h)))
  (h/2)*l+(tau-0.5)*x
  }
)
loss_uniform <- lapply(tau_vec, function(tau){
  l = ifelse(abs(x/h)<= 1,((x/h)^2/2+0.5),abs(x/h))
  (h/2)*l+(tau-0.5)*x
  }
)


df <- data.frame("Loss" = c(unlist(loss_gaussian),
                            unlist(loss_uniform)),
                 "Quantile" = rep(rep(c("0.25","0.5","0.75"),
                                      each = length(x)),2),
                 "x" = rep(x,6),
                 "Kernel" = rep(c("Gaussian", "Uniform"),
                                each = 3*length(x)))

ggplot(data = df, aes(x = x, y = Loss, color = Quantile)) + 
  geom_line() + 
  facet_wrap(. ~ Kernel) + 
  geom_hline(yintercept = 0)
```

The difference to the desired loss function $\rho_\tau$ is the largest at the origin.

```{r fig_loss_diff, echo=FALSE}

loss_diff_gaussian <- lapply(1:length(tau_vec), function(index) {
    loss_gaussian[[index]] - loss[[index]]
})
loss_diff_uniform <- lapply(1:length(tau_vec), function(index) {
    loss_uniform[[index]] - loss[[index]]
})

df$Loss_diff <- c(unlist(loss_diff_gaussian),unlist(loss_diff_uniform))

ggplot(data = df, aes(x = x, y = Loss_diff, color = Quantile)) + 
  geom_line() + 
  facet_wrap(. ~ Kernel) + 
  geom_hline(yintercept = 0) +
  ylab("Loss Difference")

```
In the figure one can see that the difference behaves uniformly over different quantiles.

The quantile boosting algorithm applies greedy algorithms to minimize the corresponding loss functions. 

## Quantile Boosting: Example
We generate data from a basic linear model 
$$ Y = \beta_0 + X^T\beta +\epsilon,$$
where $\beta_0=1$ and $\epsilon\sim t_2$. To account for a high-dimensional setting, we only generate $n = 200$ observations, whereas the covariates $X$ are generated from a standard multivariate gaussian distribution of dimension $p = 500$. The coefficient vector is set to 
$$ \beta_j = \begin{cases} 1,\quad j = 1,\dots s\\
0,\quad j = s+1,\dots,p\end{cases},$$
such that the first $s$ components are relevant. 

```{r dgp}
library(MASS)
set.seed(42)
n <- 200; p <- 500; s <- 4
beta <- rep(c(1,0),c(s+1,p-s))

X = mvrnorm(n, rep(0, p), diag(p))
epsilon = rt(n, 2)
Y = cbind(1, X) %*% beta + epsilon

n_test <- 1000
X_test = mvrnorm(n_test, rep(0, p), diag(p))
epsilon_test = rt(n_test, 2)
Y_test = cbind(1, X_test) %*% beta + epsilon_test
```

### Orthogonal Quantile Boosting/ Weak Chebyshev Greedy Algorithm (WCGA)

We start with the orthogonal variant of the qboost algorithm (WCGA, by setting `stepsize = NULL`) and proceed for $100$ greedy selection steps.

```{r example}
library(qboost)
n_steps <- 100; tau <- .5
model_WCGA <- qboost(X,Y, tau = tau, m_stop = n_steps, h = 0.2, kernel = "Gaussian", stepsize = NULL)
# selected covariates
print(model_WCGA$selection_path[1:10])
# predict values
print(head(predict(model_WCGA, newdata = X_test)))
```

At first let us take a look at the Loss
```{r fig1}
library(ggplot2)
autoplot(model_WCGA, new_Y = Y, newdata = X)
```

Of course, the loss is much more interesting on a test set.
```{r fig2}
autoplot(model_WCGA, new_Y = Y_test, newdata = X_test)
```

### Quantile Boosting / Weak Greedy Algorithm (WGA, shortened)

Additionally, we can employ the Weak Greedy Algorithm (here with $400$ steps).
```{r example2}
n_steps_WGA <- 400
model_WGA <- qboost(X,Y, tau = tau, m_stop = n_steps_WGA, h = 0.2, kernel = "Gaussian", stepsize = 0.1)
```

Again, the loss is minimized
```{r fig3}
autoplot(model_WGA, new_Y = Y, newdata = X)
```

but we have to find the right stopping criterion to minimize the out of sample loss.
```{r fig4}
autoplot(model_WGA, new_Y = Y_test, newdata = X_test)
```

### Crossvalidation

To have a reasonable stopping criterion, a crossvalidated version of the algorithms is implemented. At first, lets take a look at the orthogonal variant

```{r fig5}
cv_model_WCGA <- cv_qboost(X,Y, tau = tau, m_stop = n_steps,
                           h = 0.2, kernel = "Gaussian", stepsize = NULL)
autoplot(cv_model_WCGA, new_Y = Y_test, newdata = X_test)
```

Next, repeat the same for the non-orthogonal algorithm.
```{r fig6}
cv_model_WGA <- cv_qboost(X,Y, tau = tau, m_stop = n_steps_WGA,
                           h = 0.2, kernel = "Gaussian", stepsize = 0.1)
autoplot(cv_model_WGA, new_Y = Y_test, newdata = X_test)
```

To compare the performance to the oracle model (here in green), we apply the conquer algorithm to the first $s$ components of $X$ and report the loss on the test set.

```{r exampleplot, warning=FALSE}
library(conquer)
fit_conquer <- conquer(X[,1:s], Y, tau = tau, h = 0.2, kernel = "Gaussian")
resid_conquer <- Y_test - cbind(1,X_test[,1:s])%*%fit_conquer$coeff
loss_conquer <- smooth_check_loss(resid_conquer,tau = tau, kernel = NULL)
```

```{r, echo = F, warning=FALSE}
resid_WCGA <- Y_test - Matrix::t(cv_model_WCGA$coeff_path[1,] + Matrix::t(X_test %*% cv_model_WCGA$coeff_path[-1,]))
resid_WGA <- Y_test - Matrix::t(cv_model_WGA$coeff_path[1,] + Matrix::t(X_test %*% cv_model_WGA$coeff_path[-1,]))

loss_WCGA <- apply(resid_WCGA, 2, function(x) smooth_check_loss(x,tau = tau, kernel = NULL))
loss_WGA <- apply(resid_WGA, 2, function(x) smooth_check_loss(x,tau = tau, kernel = NULL))

df_plot <- data.frame("Loss" = c(rep(loss_conquer,n_steps_WGA+1),loss_WCGA,rep(NA,n_steps_WGA-n_steps),loss_WGA),
                      "Step" = rep(0:n_steps_WGA,3),
                      "Type" = rep(c("Oracle","WCGA", "WGA"), each = n_steps_WGA + 1))

ggplot(data = df_plot, aes(x = Step, y = Loss, color = Type)) +
  geom_line() +
  theme(legend.position = "bottom") +
  scale_colour_manual(values = c("Forestgreen","red","blue")) +
  geom_vline(aes(xintercept = cv_model_WCGA$cv_m_stop),color = "red",
               linetype = "dashed") + 
  geom_vline(aes(xintercept = cv_model_WGA$cv_m_stop),color = "blue",
               linetype = "dashed")

```



## References
* He, Xuming, et al. "Smoothed quantile regression with large-scale inference." arXiv preprint arXiv:2012.05187 (2020). [Paper](https://arxiv.org/abs/2012.05187)

* Belloni, Alexandre, and Victor Chernozhukov. "ℓ1-penalized quantile regression in high-dimensional sparse models." The Annals of Statistics 39.1 (2011): 82-130. [Paper](https://projecteuclid.org/journals/annals-of-statistics/volume-39/issue-1/%e2%84%931-penalized-quantile-regression-in-high-dimensional-sparse-models/10.1214/10-AOS827.full)

* Fernandes, Marcelo, Emmanuel Guerre, and Eduardo Horta. "Smoothing quantile regressions." Journal of Business & Economic Statistics 39.1 (2021): 338-357. [Paper](https://www.tandfonline.com/doi/full/10.1080/07350015.2019.1660177?casa_token=bkJ73Q8oXYIAAAAA%3A0g8P9Bb5elGlCBU8_bsuN_oLauFgMfQcojZmI4ERJO1WVD1M5UOw1RLRix7mMxDMMWukfZR-sbE)

